# Data transformation

## Library Package
```{r echo=TRUE}
library(dplyr)
library(readr)
```


## health data

Firstly, we read the health data from cms government website.
```{r echo=TRUE}
read_health_data <- function() {
  url = 'https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/NationalHealthExpendData/Downloads/resident-state-estimates.zip'
  temp <- tempfile()
  temp2 <- tempfile()
  # download the zipfile 
  download.file(url, temp)
  # unzip the file 
  unzip(zipfile = temp, exdir = temp2)
  # read_csv, only select the state and 2014 column 
  health_per_capita <- read_csv(file.path(temp2, "US_PER_CAPITA14.CSV")) %>% 
    filter(`Item` == 'Personal Health Care ($)') %>% 
    select('State_Name', 'Y2014') %>%
    na.omit()
  names(health_per_capita) = c("state", "2014")
  return(health_per_capita)
}

health_data <- read_health_data()
head(health_data)
```

## covid data

Secondly, we read the covid_19 data before November 30, 2020 and change the column name to make it more readable.  
```{r echo=TRUE}
url = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"

confirmed_df = read.csv(paste0(url, "time_series_covid19_confirmed_US.csv"))
# colnames(confirmed_df) = gsub("X","", colnames(confirmed_df))
death_df = read.csv(paste0(url, "time_series_covid19_deaths_US.csv"))
# colnames(death_df) = gsub("X","", colnames(death_df))

# function for transform covid time series data
transform_confirmed <- function(df) {
  info_col = df[,c("UID","Admin2","Province_State","Lat","Long_")]
  Nov_30 = df["X11.30.20"]
  df = cbind(info_col, Nov_30)
  # rename columns 
  ColName = c("UID","county","state","lat", "long", paste0('confirmed', "_Nov_30"))
  for (i in 1:6){
    names(df)[i] = paste(ColName[i])
  }
  return(df)
}

# population column is in the death time series data, use another method 
transform_death <- function(df) {
  info_col = df[,c("UID","Population")]
  Nov_30 = df["X11.30.20"]
  df = cbind(info_col, Nov_30)
  # rename columns 
  ColName = c("UID","population", paste0('death', "_Nov_30"))
  for (i in 1:3){
    names(df)[i] = paste(ColName[i])
  }
  return(df)
}

# transform the two dataframe for Nov 30th only
confirmed_df_Nov_30 = transform_confirmed(confirmed_df)
death_df_Nov_30 = transform_death(death_df)


# merge confirmed and death by UID
covid_merged = merge(confirmed_df_Nov_30, death_df_Nov_30, by = "UID") # check:3340 rows, 8 columns

# add confirm_rate column
covid_merged$confirmed_rate = covid_merged$confirmed_Nov_30 / covid_merged$population 
# get death_rate: death / confirmed 
#covid_merged$death_Nov_30 = as.numeric(covid_merged$death_Nov_30)
covid_merged$death_rate = ifelse(covid_merged$confirmed_Nov_30>0, 
                                 covid_merged$death_Nov_30/covid_merged$confirmed_Nov_30,
                                 0)

# if population == 0 remove this row
covid_merged = covid_merged[covid_merged['population']!=0,]
head(covid_merged)
```

We want to find the relationship between covid_19 and election through time. So, we generate new dataframe confirmed_d3 and confirmed_TS.

```{r echo = TRUE}
transform_confirmed_d3 <- function(df) {
  info_col = df[,c("UID","Admin2","Province_State","Lat","Long_")]
  dates = df[,c("X3.30.20","X5.30.20","X7.30.20","X9.30.20","X11.30.20")]
  df = cbind(info_col, dates)
  # rename columns 
  ColName = c("UID","county","state","lat", "long")
  for (i in 1:5){
    names(df)[i] = paste(ColName[i])
  }
  return(df)
}

confirmed_d3 = transform_confirmed_d3(confirmed_df)
subdeath = subset(death_df,select = c("UID","Population"))
confirmed_d3 = merge(confirmed_d3,subdeath, by="UID")

# cr means confirmed rate 
confirmed_d3$cr3 = confirmed_d3$X3.30.20/confirmed_d3$Population
confirmed_d3$cr5 = confirmed_d3$X5.30.20/confirmed_d3$Population
confirmed_d3$cr7 = confirmed_d3$X7.30.20/confirmed_d3$Population
confirmed_d3$cr9 = confirmed_d3$X9.30.20/confirmed_d3$Population
confirmed_d3$cr11 = confirmed_d3$X11.30.20/confirmed_d3$Population
head(confirmed_d3)
```
```{r echo = TRUE}
confirmed_TS <- confirmed_df %>% select(Combined_Key,UID,Province_State, c(`X1.22.20`:`X11.30.20`)) 
death_TS <- death_df %>% select(Combined_Key, UID,Province_State, Population, c(`X1.22.20`:`X11.30.20`))
confirmed_TS_merge = merge(confirmed_df, death_df%>%select(UID, Population), by = "UID")
confirmed_TS_merge = confirmed_TS_merge[confirmed_TS_merge['Population'] != 0, ]
head(confirmed_TS)
head(death_TS)
```



## election data

The last dataset we want to consider is the 2020 election data in USA. 

```{r echo = TRUE}
url_election = "https://raw.githubusercontent.com/JessieYee/EDAV_Covid19_Research/main/election2020.csv"

election = read.csv(url_election)
#election$Combined_Key = paste0(election$County, ", ", election$State, ", US")
election[] <- lapply(election, gsub, pattern=',', replacement='')
election["state"]<-lapply(election["state"], gsub, pattern='-', replacement=' ')
election$Biden = as.numeric(election$Biden)
election$Trump = as.numeric(election$Trump)
election$Biden_support_rate = election$Biden / (election$Biden+election$Trump)
head(election)
```

We generate new variable: Biden_support_rate to calculate the support rate of Biden. For Biden_support_rate>0.5, we can conclude that more than half of people in this state support Biden. For Biden_support_rate<0.5, we can conclude that more than half of people in this state support Trump. 


## merge data

1.  left join the covid_19 data and election

```{r echo=TRUE}
covid_merge_all <- merge(covid_merged, election, by=c("state","county"),all.x = TRUE,all.y = FALSE)

covid_state = covid_merge_all %>% 
  group_by(`state`) %>% 
  summarize(`confirmed_Nov_30` = sum(confirmed_Nov_30),
            `death_Nov_30` = sum(death_Nov_30),
            `population` = sum(population),
            `confirmed_rate` =`confirmed_Nov_30`/ `population`,
            `death_rate` = ifelse('confirmed_Nov_30'>0, 
                                  `death_Nov_30`/`confirmed_Nov_30`,
                                  0),
            `Biden` = sum(Biden,na.rm=TRUE),
            `Trump` = sum(Trump,na.rm=TRUE),
            `Biden Support Rate` = `Biden`/(`Biden` + `Trump`),
  )

covid_state <- covid_state %>% 
  mutate(`Support Which Party` = case_when(covid_state$`Biden Support Rate` > 0.5 ~ "Democratic", 
                                           covid_state$`Biden Support Rate` < 0.5 ~ "Republican"
  ))
covid_state = merge(covid_state, health_data, by = "state")
head(covid_state)
```
We separate the votes to two groups: support Democratic Party and support Republican Party by checking the `Biden Support Rate`.

2. Merge election data with confirmed_d3 dataset

```{r echo = TRUE}
confirmed_d3_all <- merge(confirmed_d3, election, by=c("state","county"),all.x = TRUE,all.y = FALSE)

confirmed_d3 = confirmed_d3_all %>% 
  group_by(`state`) %>% 
  summarize(`confirmed_Nov_30` = sum(X11.30.20),
            `confirmed_Sep_30` = sum(X9.30.20),
            `confirmed_July_30` = sum(X7.30.20),
            `confirmed_May_30` = sum(X5.30.20),
            `confirmed_Mar_30` = sum(X3.30.20),
            `population` = sum(Population),
            `Nov_rate` =`confirmed_Nov_30`/ `population`,
            `Sep_rate` =`confirmed_Sep_30`/ `population`,
            `July_rate` =`confirmed_July_30`/ `population`,
            `May_rate` =`confirmed_May_30`/ `population`,
            `March_rate` =`confirmed_Mar_30`/ `population`,
            `Biden` = sum(Biden,na.rm=TRUE),
            `Trump` = sum(Trump,na.rm=TRUE),
            `Biden Support Rate` = `Biden`/(`Biden` + `Trump`),
  )

confirmed_d3 <- confirmed_d3 %>% 
  mutate(`Support Which Party` = case_when(confirmed_d3$`Biden Support Rate` > 0.5 ~ "#407abb", 
                                           confirmed_d3$`Biden Support Rate` < 0.5 ~ "#c03735"
  ))
confirmed_d3 = merge(confirmed_d3, health_data, by = "state")

confirmed_d3$state = state.abb[match(confirmed_d3$state,state.name)]
head(confirmed_d3)
```

2. Merge new data frame: death_per_100 with covid_state data
```{r echo = TRUE}
death_per_100 = covid_merge_all %>% 
  group_by(`state`) %>% 
  summarize(
    `death_Nov_30` = sum(death_Nov_30),
    `population` = sum(population),
    `death_per_100` = `death_Nov_30`/`population`*100) %>%
  select(`state`, `death_per_100`)
covid_state = merge(covid_state, death_per_100, by='state')
covid_state$death_per_1M = covid_state$death_per_100 * 10000
head(covid_state)
```

We create a new column: death per 100 by calculating the death case divided by population.

```{r}
saveRDS(covid_state, file = "covid_state")
saveRDS(election, file = "election")
saveRDS(confirmed_d3,file = "confirmed_d3")
saveRDS(confirmed_TS, file = "confirmed_TS")
saveRDS(death_TS, file="death_TS")
saveRDS(covid_merge_all, file="covid_merge_all")
saveRDS(confirmed_TS_merge,file = "confirmed_TS_merge")
```