# Missing values

```{r include=FALSE}
library(dplyr)
```

## Checking missing value

We want to check if there is any missing value in our covid_state and confirmed_d3 data set.
```{r}
covid_state = readRDS("covid_state")
covid_state[which(is.na(covid_state$`Biden Support Rate`)),]
confirmed_d3 = readRDS("confirmed_d3")
confirmed_d3[which(is.na(confirmed_d3$`Biden Support Rate`)),]
```

We can find that Alaska and D.C have NA value for `support biden rate` variable. The county name for election data in D.C and Alaska are different from the covid_19 data. So we might loosing data when merging them together. 

## Fixing the NA in dataset

Because we want to know the election status for each state without any NA, we decide to calculate the support rate for Alaska and D.C by adding the data in election.

```{r}
election = readRDS("election")
Alaska_Biden = sum(election[election$state =="Alaska",]$Biden)
Alaska_Trump = sum(election[election$state =="Alaska",]$Trump)
DC_Biden = sum(election[election$state =="District Of Columbia",]$Biden)
DC_Trump = sum(election[election$state =="District Of Columbia",]$Trump)

covid_state[2,7] = Alaska_Biden
covid_state[2,8] = Alaska_Trump
covid_state[2,9] = Alaska_Biden/(Alaska_Biden+Alaska_Trump)
covid_state[2,10] = case_when(covid_state[2,9] > 0.5 ~ "Democratic", 
                              covid_state[2,9]< 0.5 ~ "Republican")
covid_state[9,7] = DC_Biden
covid_state[9,8] = DC_Trump
covid_state[9,9] = DC_Biden/(DC_Biden+DC_Trump)
covid_state[9,10] = case_when(covid_state[9,9] > 0.5 ~ "Democratic", 
                              covid_state[9,9]< 0.5 ~ "Republican")
head(covid_state)

confirmed_d3[2,13] = Alaska_Biden
confirmed_d3[2,14] = Alaska_Trump
confirmed_d3[2,15] = Alaska_Biden/(Alaska_Biden+Alaska_Trump)
confirmed_d3[2,16] = case_when(confirmed_d3[2,15] > 0.5 ~ "Democratic", 
                              confirmed_d3[2,15]< 0.5 ~ "Republican")

confirmed_d3[9,13] = DC_Biden
confirmed_d3[9,14] = DC_Trump
confirmed_d3[9,15] = DC_Biden/(DC_Biden+DC_Trump)
confirmed_d3[9,16] = case_when(confirmed_d3[9,15] > 0.5 ~ "Democratic", 
                              confirmed_d3[9,15]< 0.5 ~ "Republican")



```

```{r}
saveRDS(covid_state,file = "covid_state")
saveRDS(confirmed_d3,file = "confirmed_d3")
```



